---
title: 'Stat 133 HW01: Exploratory Data Analysis and Getting Familiar with R Expressions'
author: "Skyler Roh 23313980"
output: html_document
---

### Introduction

This assignment has two purposes: 
(a) to familiarize you with some of the basic 
functions in R and how to use them, and 
(b) to review/introduce some of the key
concepts in exploratory data analysis.

We will explore a small data set collected by Kleiber. 
He wanted to study the relationship between an animal's metabolic
rate and mass.

Kleiber measured the metabolic rate and mass of 19 animals 
(mouse, dog, cat, goat, man, cow, elephant, ...)

We begin by reading the data from the Web into R.

```{r}
load(url("http://www.stat.berkeley.edu/users/nolan/data/Kleiber.rda"))
```

We have loaded several R objects, from the rda file on the Web. 
The following functions may be useful in answering 
the questions:
**class(), cor(), density(), length(), ls(), mean(), min(),
max(), plot(), quantile(), summary().**


1. What are the names of the objects?
What type of information are they, i.e., are they numeric, logical,
character, factors? 

```{r}
ls()
name
class(name)
```

2. Can you confirm that there are measurments on 19 animals 
available in the Kleiber rda file? (Use the length() function).

```{r}
length(name)
```


The data do not indicate the units of measurement. 
metabolic rate is in kilo-calories per day, i.e. kcal/day, and
mass is in kilograms.

We are ultimately interested in the relationship between mass and 
metabolic rate, but it's a good idea to begin an exploration of the
single variables first. 

### Metabolic Rate

We wish to summarize the distribution of the values for metabolic rate. 
We can summaarize them numerically or graphically.

We begin with numeric summaries. 
We often summarize values by providing a measure of
the center or middle value and the typical spread of 
the values about the center.

3. One measure of center is the mean or average.
What is the average metabolic rate?  

```{r}
mean(mr)
```


4. Read the documentation for the mean function and find the 
10% trimmed mean and the 50% trimmed mean of metabolic rate. 
A trimmed mean removes the extreme values before taking the average.  
For example, a 10% trimmed mean drops the smallest 10% and the 
largest 10% of the values and takes the mean of the 
remaining 80% of the data.


```{r}
mean(mr, trim = .1)
mean(mr, trim = .5)
median(mr)
```

5. How do these versions of the mean compare to each other and to 
the median metabolic rate? 
#10% trim mean >> median (removed extreme low end values which were much farther from the mean)
#50% trim mean == median

6. The measure of spread associated with the mean is the standard
deviation, or SD for short.
And the measure of spread associated with the median is the 
inner quartile range, or IQR for short.

Use Google to find the functions in R that compute these quantities
and report the SD and IQR for metabolic rate.
Save these quantities as mrSD and mrIQR, respectively.

```{r}
mrSD = sd(mr)
mrIQR = IQR(mr)
```


The numeric summaries are not as informative as graphical summaries 
of the distribution. 
With a histogram or density curve, we can get a better sense of how
the observations are distributed across the values.
We typically look at particular features of the distribution,
such as the symmetry or skewness; the number and location of modes;
the length of the tails; and any gaps where there are no observed values
or abnormally large/small values.

7. Make a density curve of metabolic rate by composing the plot
function with the density function. Play around with the bw parameter
that specifies the amount of smoothness in the curve until you
have a curve that captures the shape od the data.

```{r}
plot(density(mr, bw = 500))
mr
```

Describe the distribution of metabolic rate using the criteria
mentioned above.

#greatly skewed with a long right tail
#2 notable modes, 0-500 and ~8000, and one point out at 49000
#no points between 8166 and 49000

8. With highly skewed data or data with a few large values, it 
can be difficult to see the distribution of the bulk of the data.
Additionally, many statistical techniques rely on the distribution
being roughly symmetric. 
Repeat the visual examination of the distribution for 
the log of metabolic rate.  What can you see now that was
not aparent in the earlier plot?

```{r}
plot((density(log(mr),bw = 1)))
#nearly symetric about 6, notable small peak at ~1
```


### Mass

9. Explore the distribution of the mass of the animals. 
Make a graphical summary of the distribution and describe its general shape.

```{r}
plot(density(mass, bw = 50))
#long right tail
#peak at 0-4.2
#small peak at ~450 
#extreme point at 3672
```

10. In addition, report a five-number summary of the data.
(The five-number summary consists of the min, lower quartile,
median, upper quartile, and max of the values).

```{r}
summary(mass)
```

11. Use subsetting to find the name of the animal with the
lowest metabolic rate and the animal with the highest metabolic rate.

```{r}
name[which(mass == min(mass))]
name[which(mass == max(mass))]
```


### Relationship Between Metabolic Rate and Mass

Now that we have a clearer picture of the distribution of each of
our variables of interest, we can examine how they vary together.

Again, we can make a numeric summary and a graphical summary of
the relationship.

12. For a graphical summary, make a scatter plot of metabolic rate
against mass. Use the plot() function to do this. 
Consider plotting the original values and the log of the values
for metabolic rate and mass (examine all four combinations of 
transformations of the two variables, e.g., log MR and log Mass, 
log MR and Mass, MR and log Mass, and MR and Mass).
Choose the plot that provides the most informative view of
the relationship between these two variables and provide its plot
below. 

```{r}
plot(log(mr), log(mass))
```

13. The correlation between two variables is a measure of linear 
association.  Find the correlation between the log of metabolic rate 
and the log of mass. How does this compare 
to the correlation between metabolic rate and mass?

```{r}
cor(log(mr), log(mass))
cor(mr, mass)
#cor(log(mr), log(mass)) ~< cor(mr, mass)
```

Even though the correlation is somewhat lower with the transformed
data, it's a good idea to take logs. Why? Take a look at the scatter
plot of the untransformed data and notice the influence of the largest
animal.

14. Metabolic rate is measured in kcal/day. 
Convert metabolic rate to joules per hour. 
Save these transformed values in the variable called newMR.
Then, compute the correlation between the new metabolic rate and mass. 
How does the correlation change?
The correlation coefficient is said to be unitless.
Why does this property explain the correlation between the revised
metabolic rate and mass?

```{r}
newMR = mr*1000/24
cor(newMR, mass)
#cor(newMR, mass) = cor(mr, mass)
#newMR is only a unit change, thus it does not change the correlation which is a unitless measure
```


