---
title: "Predictors of Spam"
output: html_document
author: Skyler Roh 23313980
---


```{r}
load("emailDF.rda")
colors = c("red", "green")
factorSpam = factor(emailDF$isSpam, levels = c(TRUE, FALSE), labels = c("Spam", "Ham"))
spamColors = colors[factorSpam]
```

Add your two plots and description

```{r, echo=FALSE}
plot(emailDF$percentCapitals, col = spamColors, xlab = "", ylab = "", pch = 19, cex = .25)
title(main = "Spam or Ham: \nPercentage of Capital Letters", xlab = "Index of Email", 
      ylab = "Percentage of Capital Letters in Body")
legend("topleft", title = "Spam or Ham", fill = colors, legend = levels(factorSpam), cex = .7)
```
After graphing the percentage of capital letters in the body of each email and distinguishing Spam vs Ham based on color, the right hand portion of the plot (Spam) shows a greater frequency of high percentage of capital letters (>40%). Also, there is a much higher frequency of all capital letters, indicated by the consistent band of red points at the top of the plot, amongst the Spam messages. 


```{r, echo=FALSE}
plot(log(emailDF$exclaimCount), col = spamColors, xlab = "", ylab = "", pch = 19, cex = .4)
title(main = "Spam or Ham: \nExclamation Mark Count", xlab = "Index of Email", ylab = "Log(Number of Exclamation Marks in Body)")
legend("topleft", title = "Spam or Ham", fill = colors, legend = levels(factorSpam))
```
After graphing the log of the number of exclamation marks in the body of each email and distinguishing Spam vs Ham based on color, the right hand portion of the plot (Spam) again shows a higher density of relatively high numbers of exclamation marks relative to that of spam messages. The highest indices that are Ham do show higher levels of exclamations; however, the high counts amongst Spam is much more consistent. Thus high numbers of exclamations are most likely indicate a Spam message. 
